Embedded Scalars: {A}n Experimental Revisit

### 2014-05-29 text/paper/paper4.tex
Conclusions sec com

Traditionalism, on the other hand, has to explain the possibility of
local readings of intonationally unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentence
compared to that in accented \as-sentences. To do so within a
traditionalist framework seems impossible, so the most common approach
is to explain putative evidence for local readings away as the result
of some other phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that the additional assumption of 
``shallow processing'' of the non-monotonic quantifiers might save
traditionalism.

@@

\bigskip

\paragraph{Some Potential Objections.} There are a number of
conceivable objections to our interpretation of the presented
results. We will consider the most pressing ones in the following.

@@

A third objection could be that pictures in our tasks may unduly
overemphasize the contrast between cases where items are connected
with all or merely some but not all of their surrounding shapes. This
is essentially the line of argument pursued by
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} against the data
reported by \citet{ChemlaSpector2010:Experimental-Ev} in support of
local readings of
\es-sentences. \citeauthor{GeurtsTielvan-Tiel2013:Scalar-expressi}
show that the strength of agreement to an \es-sentence on a 7-point
Likert-scale depends on a visual contrast between, on the one hand,
one element connected to some but not all of its surrounding shapes,
and, on the other hand, a variable number of elements connected to all
of their surrounding shapes. When two universally connected elements
are present, the mean rate of agreement with a suitable \es-sentence
is higher than when only one is present. So, maybe our displays
similarly provoked local readings because of an overemphasized visual
contrast.

This cannot be a plausible explanation for the number of local
responses we found in our study. Although it is true that the critical
position for judging the local reading is a picture with a high visual
contrast in the above sense (see Figure~\ref{fig:exseqES}), a purely
contrast-based explanation of the local responses in our data is not
feasible for two reasons. The first is that visual contrast does not
explain how subjects ever even got past the second critical position
where the sentence can be judged false under its literal reading. At
this position there is a visual contrast of the alleged kind but it is
as small as possible. So something must have motivated subjects
already at this position to continue unraveling, given that they were
generally able of exiting at the earliest possible location, as
evidenced by ceiling success rates in our positional controls. A second
argument that casts doubt on a contrast-based explanation is that our
study elicited categorical truth-value judgements. Unlike in
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s
(\citeyear{ChemlaSpector2010:Experimental-Ev}) and
\citeauthor{GeurtsTielvan-Tiel2013:Scalar-expressi}'s
(\citeyear{GeurtsTielvan-Tiel2013:Scalar-expressi}) experiments, we
did not record degrees of agreement with a statement. But it is hardly
plausible that the visual contrast in a picture like in
Figure~\ref{fig:subfig1} alone is enough to overrule a truth-value
judgment so as to contradict the semantic meaning (even when there is
no intonational markedness to support such a
reinterpretation). Ultimately, however, even if the local answers in
our study are contrast effects, the verdict remains that
traditionalism can not account for that piece of the data unless aided
by yet another auxiliary and grammaticalism cannot explain why it is
relatively dispreferred.

% A fourth major conceivable objection maintains that we cannot rule out
% that the way we realized the intonational markedness in the accented
% conditions was not the right kind of intonational
% markedness. Therefore the intonational markedness hypothesis that
% links an increase in local readings to intonation might not be
% falsified by our results. Although true, this still does not help
% vindicate traditionalism, for even without marked intonation local
% readings were available.

Also, we should address the obvious worry that maybe our results
could have been the product of typicality after all. Remember that
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} argue that the
distribution of responses to \as-sentences found by
\citet{CliftonDube2010:Embedded-Implic}, as well as
\citet{ChemlaSpector2010:Experimental-Ev} could be explained by
differences in how typical representations for an \as-sentence the
presented pictures were. To apply typicality-explanations to the
incremental verification task, one would have to reason about the
typicality of parts of pictures or to figure in subjects' expectations
of pictorial typicality given their uncertainty about parts of the
picture that they have not yet seen. On intuitive grounds, we take
either extension to be far-fetched and implausible. But even if this
line of explanation can be plausibly extended to incrementally
revealed picture verification, this does not as such save the day for
traditionalism or grammaticalism, because, as argued above, typicality
effects are arguably insufficient to explain the change in categorical
truth-value judgement that we found in our experiment.

% Still, graphical properties might affect judgments in the worst
% case. In the first step of the \as-condition local and global readings
% cannot be judged, yet. Therefore, graphical properties should not
% affect judgments if participants have one of these readings in
% mind. If participants have a literal reading in mind, the worst case
% scenario is that they refrain from judging the sentence as true at
% this point due to some, say, atypical, graphical property of the
% picture. The situation is similar at the second step. The local
% reading is still not decidable. Therefore, graphical properties should
% not play a role if participants have a local reading in mind. A
% positive judgment might, however, be a delayed judgment of a literal
% reading. Again, the worst case scenario for our purpose is that
% participants refrain from answering because of graphical
% properties. At the last step the local reading requires a negative
% judgment. Now, we could obtain delayed negative responses from literal
% or global readings because the graph as a whole is, e.g., logically
% true but atypical with respect to these readings. These kinds of
% judgments would be indistinguishable from local responses. Note,
% however, that we have no reason to believe that typicality has large
% effects, because, given the results from van Tiel, the picture as a
% whole is expected to be one of the more typical situations for
% \as-sentences. Still, we cannot exclude the possibility that graphical
% properties of the pictures might have the described effect. In the
% \es-conditions this type of effect is, however, entirely impossible
% since the literal and global reading is plainly false given the
% initial parts as well as the whole graph. Summing up, in the worst
% case, which we consider implausible, graphical effects might lead to
% an enhanced number of local responses in the \as- as compared to the
% \es-condition. We have to keep this possibility in mind. We want to
% stress, however, that the design presented here improves on previous
% studies with respect to possible graphical effects.

>>>>

Traditionalism, on the other hand, has to explain the possibility of
local readings of intonationally unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentence
compared to that in accented \as-sentences. To do so within a
traditionalist framework seems impossible, so the most common approach
is to explain putative evidence for local readings away as the result
of some other phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that the additional assumption of 
``shallow interpretation'' of the non-monotonic quantifiers might save
traditionalism.

@@

A more promising alternative explanation for local readings of
\es-sentences is pursued by
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} \citep[see
also][]{Tielvan-Tiel2014:Quantity-Matter} in response to the data
reported by \citet{ChemlaSpector2010:Experimental-Ev}. The main idea
is that visual contrast between an item that is connected to all, and
one that is connected to only some but not all of the relevant
elements may trigger exceptional local enrichment of \emph{some}. In
support of this idea,
\citeauthor{GeurtsTielvan-Tiel2013:Scalar-expressi} show that the
strength of agreement to an \es-sentence on a 7-point Likert-scale
depends on how strong the relevant visual contrast is. When an item
with only some connections is contrasted presented alongside two
universally connected elements, the mean rate of agreement with a
suitable \es-sentence is significantly higher than when only one
universally connected element is present. So, maybe our \es-sequences
similarly provoked local readings because of an overemphasized visual
contrast between \emph{some but not all} and \emph{all}.

The third step of our \es-sequences in Figure~\ref{fig:exseqES} is a
relevant case of direct contrast between a \emph{some-but-not-all-}
and an \emph{all}-situation, but it is a weak contrast, in the sense
of \citet{GeurtsTielvan-Tiel2013:Embedded-Scalar}, because there is
only one universally connected element. Moreover, our study elicited
categorical truth-value judgements. Unlike in
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s
(\citeyear{ChemlaSpector2010:Experimental-Ev}) and
\citeauthor{GeurtsTielvan-Tiel2013:Scalar-expressi}'s
(\citeyear{GeurtsTielvan-Tiel2013:Scalar-expressi}) experiments, we
did not record degrees of agreement with a statement. For a purely
contrast-based explanation to work, it would have to be made plausible
why the (weak) visual contrast in a picture like in
Figure~\ref{fig:exseqES3} alone is enough to overrule a truth-value
judgment so as to contradict the semantic meaning (even when there is
no intonational markedness to support such a reinterpretation).

What might make this contrast-based explanation more plausible is the
observation that subjects found \es-sentences difficult. As discussed
in Section~\ref{sec:results}, there were a non-negligible number of
\emph{true} answer at step 2 in the \es-sequence. This can be
explained on the assumption that participants read the non-monotonic
quantifier \emph{exactly one} as a monotonic quantifier \emph{at least
  one}. Similar ``shallow interpretation'' of \emph{exactly one} would
explain the high number of local answers for \es-sentences (and for
\es-sentence alone). If participants processed the truth-value for an
\es-sentence by looking for a \emph{unique} element that could be
singled out somehow, they would need to unravel the \es-sequence up to
the last picture. At this step contrast effects are stronger and would
not need to compete with the proper semantic interpretation of
\emph{exactly one}. Together, pictorial contrast effects in symbiosis
with a ``shallow interpretation'' non-monotonic quantifiers could be
an alternative explanation of (\ref{bsp:observed-preferences}) for
traditionalism.

\paragraph{Critique.} There are a number of conceivable objections to
our design and our interpretation of the presented results. We will
consider the most pressing ones in the following.

@@
